### 2023-08-09 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

\newcommand{\Tcool}{\input{output/T1.tex}}
\newcommand{\Thot}{\input{output/T2.tex}}
\newcommand{\eom}{\input{output/om_flare.tex}}
\newcommand{\eepic}{\input{output/epic_flare.tex}}
\newcommand{\Tqmean}{\input{output/Tqmean.tex}}
\newcommand{\Tfmean}{\input{output/Tfmean.tex}}
\newcommand{\FX}{\input{output/epic_flux.tex}}
\newcommand{\LX}{\input{output/epic_Lx.tex}}
\newcommand{\Lbol}{\input{output/Lbol.tex}}
\newcommand{\LXLbol}{\input{output/lxlbol}}
\newcommand{\ffdalpha}{\input{output/tess_ffd_alpha}}
\newcommand{\ffdbeta}{\input{output/tess_ffd_beta}}
\newcommand{\ffdr}{\input{output/R315}}

% Title
\title{Fully convective star that produced a near-polar flare shows a typical corona}

% Author list with  \affiliation

\author[0000-0002-6299-7542]{Ekaterina Ilin}
\affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}

\author[0000-0003-1231-2194]{Katja Poppenh\"ager}
\affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}
\affiliation{Institute for Physics and Astronomy, University of Potsdam, Karl-Liebknecht-Strasse 24/25, 14476 Potsdam, Germany}

\author{Beate Stelzer}
\affiliation{Institut für Astronomie \& Astrophysik, Eberhard-Karls-Universität Tübingen, Sand 1, 72076 Tübingen, Germany}
\affiliation{INAF – Osservatorio Astronomico di Palermo, Piazza del Parlamento 1, 90134 Palermo, Italy}

% Abstract with filler text
\begin{abstract}
    In 2020, TESS observed a rapidly rotating M7 dwarf produce a flare at 81 deg latitude, the highest latitude flare located to date. This is in stark contrast to solar flares that occur much closer to the equator, typically below 30 deg. The circumstances that allow flares at high latitudes to occur are poorly understood. We studied five Sectors of TESS observations, and obtained 36 ks of XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to X-ray luminosity, coronal temperatures, and flaring activity. Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band. Combining the 18 flares found in TESS with the flare from XMM-Newton, we inferred that TIC 277 produces flares at or above this energy approximately once every 10 h. We conclude that TIC 277's X-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.

%    and determined that the X-ray luminosity $L_X$ is slightly lower but not far below the $L_X$ of other stars with similar spectral types a rotation rates. We constructed flare frequency distributions using the three TESS Sectors, with a power law slope of $XXXX\pm XXXX$, and flare rate above $\log_{10}E=31.5\,$erg of XXXX, again typical of similar stars. Our results suggest that the star's magnetic activity is typical for its rotation rate and mass. However, the rotation period of 4.56 hours is on the short end, which could indicate that polar flares on these stars are a feature of youth.

\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}
Most M dwarfs hosts terrestrial planets with a sizable fraction orbiting at instellations where liquid water can exist~\citep{dressing2015occurrence, hardegree-ullman2019kepler, ment2023occurrence}. However, to be habitable in an Earth-like manner, merely the possibility of surface water in its liquid form does not suffice. Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere affects hospitability for life, for M dwarfs in particular~\citep{airapetian2020impact}. If the energetic (particle) flux is too high, the atmosphere may blow off~(XXXX). If not energetic enough, life may not emerge in the first place~\citep{rimmer2018origin}.

In M dwarfs, stellar activity evolution on the main sequence unfolds much slower than for a FGK star. While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{XXXX}. Moreover, the zone where liquid water can reside is much closer to these faint stars. The result is a planet that may be exposed to violent space weather conditions for a good fraction of the age of the universe.

Stellar space weather originates from the stellar corona. The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission. The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~(XXXX). Yet, brown dwarfs are regularly detected with energetic flares regardless of their low X-ray luminosity~(XXXX). How these flares are produced in the apparent absence of a solar-type corona is unclear~\citep{weber2023understanding}.%. An intermittent corona is conceivable, similarly XXXX. Perhaps, the rapid rotation of these low-mass objects plays a role, as it governs the dynamo process within, and consequently influences the complex process of flux emergence

\citet{ilin2021giant} searched fully convective dwarfs that straddle the transition from star to brown dwarf with TESS. They found four flares on four rapidly rotating ($P<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star. The flares were observed to rotate in and out view. The shape of this modulation together with known inclination (combining $P$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares. Surprisingly, these flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below 30 deg latitude. Their results indicate a preference for flares at high latitudes in these stars, as a chance finding was unlikely ($\sim 0.1\%$). If flares and associated particle eruptions indeed have a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought. %Are high flaring latitudes  an effect of the dynamo at high rotation speeds? Or is there something special about these stars' coronae that allowed them to produce flares at these latitudes?

%Planets around late M dwarfs exists, see TRAPPIST-1 or LHS 3154~\citep{stefansson2023extreme}

TIC 277539431, TIC 277 for short, is one out of four stars with a high-latitude flare in its TESS observations~\citep[][Table \ref{tab:starparams}]{ilin2021giant}. In fact, it showed the highest latitude flare known to date at $\sim81\,$deg. TIC 277 is an M7 dwarf with an extremely short rotation period of 4.56 hours, detected in the rotational modulation of its light curve. While the fast rotation and late spectral type could be indicative of a brown dwarf~(), the detection of flares suggests the presence of a stellar corona~().%The star's age is unknown. \citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would, if not as high, be compatible with the NUV/FUV emission from younger open cluster members. 
%Is its high flaring latitude an effect of the dynamo at high rotation speeds? Or is there something special about this star's corona that led it to produce are flare at these latitudes? 
In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from the Transiting Exoplanet Survey Satellite~(TESS,~\citealt{ricker2015transiting}, Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}). We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star, brown dwarf, or something in between~(Section~\ref{sec:discussion}), and conclude in Section~\ref{sec:summary}.

%producing a high-latitude flare requires an unusual corona, or if it is rather typical of stars of this spectral type and rotation.

\begin{table}
% \footnotesize
\centering
    \caption{TIC 277539431 (WISEA J105515.71-735611.3)}
    \begin{tabular}{ll}\hline 
         Parameter & Value  \\\hline
         Distance $d$ & $13.70\pm0.11\,$pc [2] \\
         % $G$ mag & $14.7189 \pm 0.0011$ [4]\\
         % $G_{BP}$ mag & \\
         % $G_{BP}$ mag & \\
         % $K_s$ mag & $9.666 \pm 0.024$ [3]\\
         % $J$ mag & $10.630 \pm 0.023$ [3] \\
         Spectral type & M7 [1]\\
         Effective temperature $T_{\rm eff}$ & $2680^{60}_{-50}\,$ K [3]\\
         % age & \\
         Rotation period $P_{\rm rot}$ & $273.618 \pm 0.007\,$min [1]\\
         Projected rotation velocity $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
         Inclination $i$ & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
         Radius $R$ & $0.145\pm0.004\,R_\odot$ [1]\\
         Bol. luminosity $L_{\rm bol}$ & \Lbol \\\hline

    \end{tabular}
    \newline\footnotesize
    [1] \citet{ilin2021giant}, 
    [2] \citet{bailer-jones2018estimating}, 
    % [3] 2MASS, \citet{skrutskie2006two}, 
    % [4] \citet{collaboration2016gaia,gaiacollaboration2023gaia}, 
    [3] \citet{pecaut2013intrinsic}
    \label{tab:starparams}
\end{table}

\section{Observations}
\label{sec:obs}

\subsection{XMM-Newton}
\label{sec:obs:xmm}
XMM-Newton is an X-ray telescope, which launched into orbit 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
We used XMM-Newton to observe TIC 277 for 36 ks, i.e. two full rotation periods of the star, on August 5, 2022 (Proposal 090120; PI: E. Ilin). We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band between 0.2-12 keV, as well as the OM using its broad band white light filter.

\subsubsection{Time series with EPIC and OM}

 We extracted the events time series from EPIC using XMM SAS version 20 using the \texttt{evselect} task. We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively. We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extract a soft X-ray time series with a time binning of 200 s.

Simultaneously, we used the optical monitor (OM) aboard XMM-Newton, to monitor TIC 277 in its white light filter at $10$\,s cadence. The OM white light filter is centered on $4,060\AA$, with an equivalent width of $3,470\AA$. We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.

Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve.

\begin{figure*}[ht!]
    \script{FIGURE_lightcurves.py}
    \begin{centering}
        \includegraphics[width=0.75\linewidth]{figures/lightcurves.png}
        \caption{
         Top panel: Optical Monitoring (OM) light curve. Bottom panel: X-ray (PN, MOS1 and MOS2 combined) light curve. The grey shaded portion defines the flare-only subset of the observations (see Section~\ref{sec:methods:epic} and Table~\ref{tab:specfit}).
        }
        \label{fig:lightcurves}
    \end{centering}
\end{figure*}

\subsubsection{EPIC Spectra}
From the EPIC observations, we used standard spectral extraction procedure for point sources. We extracted spectra using \texttt{evselect} to select all events in the circular event and background regions,\texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra. Finally, we use \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin. To prduce spectra for the quiescent and flaring portions of the observations separately, we used \texttt{tabgtigen} to select the time intervals to pass to \texttt{evselect}.

\subsection{TESS}
\label{sec:obs:tess}
Since 2018, TESS has been supplying publicly available red-optical high-precision time series photometry in an ongoing all-sky survey. Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $6,000$-$10,000\AA$ band pass.

We used 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence by TESS. Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023~(Fig.~\ref{fig:tess_lcs}).

We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}. The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and the two Savitzky-Golay filters with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking outliers above $>2.5\sigma$, as well as the six-fold of consecutive originally masked data points following the original mask, and finally, additional 3 adjacent data point before and after the so expanded mask.

In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level in the masked light curve, filling in the noise level inside the masked areas with the mean of the noise levels adjacent to the mask. To the series of outliers, we kept adding data points to the flare candidate until a data point falls below $2\sigma$.

\begin{figure*}
    \script{FIGURE_tess_lcs.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/tess_lcs.png}
        \caption{
         Normalized TESS light curves. Black and red dots show the light curve with and without rotational variability and trends, respectively. The de-trended light curve is offset by $0.2$. The large, rotationally modulated, flare, localized at about 81 deg latitude by~\citet{ilin2021giant}, appears in the second half of the top panel~(Sector 12).}
        \label{fig:tess_lcs}
    \end{centering}
\end{figure*}

\section{Methods}
\label{sec:methods}

\subsection{Spectral fitting: EPIC}
\label{sec:methods:epic}
We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}. We analyzed both the full data set, and the quiescent and flaring parts separately. Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model either. In the full data set, we used the \texttt{bayes} method in XSPEC to assign constant priors within the model's hard limits, and sample the uncertainties using the Markov-Chain-Monte-Carlo (MCMC) method using \texttt{chain} for a total of 30000 steps, discarding the first 5000 as the burn-in phase. From the same fit, but using PN data only, we derived the flux and X-ray luminosity in the 0.2-12 keV band. For the quiescent and flaring portions of the observations, we chose the same procedure, but used Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest of the upper and lower uncertainties as standard deviation.

\subsection{Flare energies: TESS, OM, and EPIC}

We fitted the soft X-ray light curve using constant flux with a flare, applying the~\citet{davenport2014kepler} optical flare model that captures the fast rise and exponential decay of the flare seen with XMM-Newton. From the fit, we extracted the baseline flux against which we measured the equivalent duration $ED$ of the flare, i.e., the flare flux $F_{\rm flare}$, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:

\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}

We then converted $ED$ to flare energy in the $0.2-2\,$keV band by multiplying with $L_X$.

For the OM flare, we derived the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}. We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}. From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:

\begin{equation}
    f_{\rm OM} = \frac{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})}  d\lambda}{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})} d\lambda} 
    \label{eq:ratio}
\end{equation}

With this, we can calculate the bolometric flare energy as:

\begin{equation}
    E_{\rm flare} = ED \cdot \pi R^2 \cdot \sigma_{kB} T_{\rm eff}^4 \cdot f_{\rm OM}
    \label{eq:eflare}
\end{equation}

We note that due to the optimistic response curve assumption, the resulting $E_{\rm flare}$ is likely higher in reality.

% To obtain the corrected $ED$, we then multiplied with $L_{\rm bol}$ to arrive at the bolometric flare energy. $L_{\rm bol}$ was derived using G and J magnitudes by converting J to V magnitude, then using the bolometric correction from \citep{mann2015how,mann2016erratum}.

Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the TESS response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the redder optical filter of TESS, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}. Besides the $ED$, \texttt{altaipony} also yields flare start and end times, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitudes $a$.

\section{Results}
\label{sec:results}
With the XMM-Newton and TESS observations, we could constrain the coronal and flaring properties of TIC 277. We derive its coronal temperature and luminosity from the EPIC spectrum~(Section~\ref{sec:res:XrayTL}). We also measure the flare energies in X-ray from the EPIC instruments, and optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in TESS~(Section~\ref{sec:res:flares}).

\subsection{Coronal temperature and luminosity}
\label{sec:res:XrayTL}
% From the proposal: $F_X=2.7^{-13}$ erg/s/cm$^2$, $L_X=6\cdot10^27$ erg/s (update with new eROSITA data) -- 1 oom above the measured XMM values.

From the XMM-Newton EPIC spectra, we could measure the coronal properties of TIC in both its quiescent, and flaring states~(Table~\ref{tab:specfit}). Using the combined quiescent and flaring data, we found a two-temperature VAPEC model to fit better than a one-temperature solution and no worse than a three-temperature solution. The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively. With these temperatures as ground truth, we also fitted two-temperature VAPEC models to the flaring and quiescent portions of the observations separately. The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall. However, the emission measure weighted mean temperature change dfrom \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum. This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona. In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).

Not using the full set of observations as ground truth resulted in an unconstrained hot coronal component. This can be interpreted as due to a lack of X-ray flux above 5 keV. Hot flares with temperatures of >50 MK have been measured on DS Tuc A~\citep{pillitteri2011xray}. On the Sun, the highest flare associated coronal temperatures are around 35 MK~\citep{kay2003soft}.

From the PN spectra, we measured a soft X-ray flux of \FX, and a luminosity \LX in the $0.2-2\,$keV band. 
 \begin{table}
% \movetableright=-20mm
\centering
    \script{TABLE_specfits.py}
    \caption{XSPEC fits to EPIC spectra for different subsets of observations.}
    \input{output/mcmc_specfit.tex}
        \label{tab:specfit}
\end{table}

\begin{figure}
    \script{FIGURE_data_resid.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/joint_chain_fit_data_resid.png}
        \caption{
         Best-fit joint Double APEC spectrum fit with Fe/0$=0.6$.
        }
        \label{fig:spec_joint_all}
    \end{centering}
\end{figure}

\subsection{Flares}
\label{sec:res:flares}

\begin{table}
% \movetableright=-20mm
\centering
    \script{TABLE_tess_flares.py}
    \caption{Flares detected with TESS. $a$ is the relative amplitude, and $E_{\rm bol}$ is the bolometric flare energy assuming a 10,000 K blackbody emission from the flare.}
    \input{output/tess_flares.tex}
        \label{tab:flares}
\end{table}

In TESS, we found a total of 18 flares in the five Sectors, or equivalently, 125 days of observing at 2-min cadence. We fit the flare frequency distribution (FFD) with a power law of the form

\begin{equation}
    f(E) \mathrm{d} E = \beta \cdot E^{-\alpha} \mathrm{d} E.
\end{equation}
following the procedure in~\citep{ilin2021flares}, implemented as \texttt{FFD.fit\_powerlaw} using the \textit{mcmc} argument in \texttt{AltaiPony}~\citep{ilin2021altaipony}, which is using the posterior distribution derived in \citet{wheatland2004bayesian}, and sampling the posterior distribution with the \texttt{emcee} package~\citep{foreman-mackey2013emcee}.

The FFD power law fit converged after 13,500 steps on a slope $\alpha =$\ffdalpha. The frequency $R_{31.5}$ of flares per day above $\log_{10} E = 31.5\,\mathrm{erg}$ is about $\log_{10}R_{31.5}=$\ffdr. The slope is typical of other flaring stars, regardless of spectral type a rotation period~(see, e.g., Fig.~13 in \citealt{ilin2021flares}). $R_{31.5}$ is typical of flaring M dwarfs, in particular typical of late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.

For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-5\,$keV band. The energy of the corresponding OM flare is \eom, which fits the extension of TESS' flare frequency distributions to lower energies within uncertainties on the power law fit~(Fig.~\ref{fig:ffd}). About one flare every $10\,$h of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277, assuming all optical flares above this energy in the OM data correspond to X-ray flares of similar or higher energy to the observed. \citet{guarcello2019simultaneous} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled energies are above $10^{33}\,$erg both in X-ray and optical. \citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result. We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely eclipsed by the star~\citep{johnstone2012soft}.

\begin{figure}
    \script{FIGURE_tess_ffd.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/277539431_tess_ffd.png}
        \caption{
         Cumulative flare frequency distribution and power law fit.
        }
        \label{fig:ffd}
    \end{centering}
\end{figure}

\section{Discussion}
\label{sec:discussion}
% \citep{pass2022constraints} on rotational evol of late Ms

Is the magnetic activity of the star with the highest known flare latitude known to date is any way unusual? Our findings indicate that TIC 277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}), X-ray luminosity~(Section~\ref{sec:discussion:xraylum}), and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of stars with its rotation rate and mass. TIC 277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}. This could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares that other dynamo states cannot generate. With TESS and XMM-Newton, we can now evaluate if this suspected exceptionality is reflected in the star's magnetic activity and coronal properties.

\subsection{Coronal temperature}
\label{sec:discussion:xraytemp}
Low mass stars' coronae are not uniform. Magnetic structures create a range of temperatures, starting at about 1 MK by definition and rising up to 30 MK in flares, in the heterogeneous Solar corona~\citep{vaiana1978recent}. Stellar point source X-ray observations are a superposition of various temperatures and densities. The resolution of different temperature component depends on the apparent brightness and integration time of the spectrum. Nearby stars like Proxima Cen can be modeled with a range of temperatures without overfitting~\citep[e.g.,][]{gudel2004flares, drake2020pointing}, while one can often only capture one dominant temperature in fainter objects~(XXXX).

We find that the quiescent corona of TIC 277 is best described by two components, similar to many other active M dwarfs~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}. We measure temperatures of 3 MK and 12 MK with an emission-measure weighted average of about 6.5 MK, which rises to 11 MK during the flare~(Table~\ref{tab:specfit}). The hotter 13 MK component represents thermal bremsstrahlung radiation from hot plasma that evaporates from the chromosphere into the coronal flaring loops~\citep{benz2016flare}, which is supported by the observed delay between the optical and the X-ray flare~\citep{hawley2003multiwavelength}. The optical flare originates from the lower chromosphere, which is heated by the precipitating particles. This heating then causes evaporation of particles into the flare loop. A third, very hot, $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV.

The range of 6.5-11 MK for the emission measure weighted temperature of TIC 277 places it squarely in the saturated regime with other mid M dwarfs down to spectral type M6~\citep{wright2018stellar, magaudda2020relation, stelzer2022first,robrade2005xray,raassen2003xray,paudel2021simultaneous, foster2020corona}. Flares on the slowly rotating Proxima Cen (M5.5, 85 d rotation~XXXX), appear with a range of temperatures at 2-3 MK and 10-20 MK~\citep{gudel2004flares, fuhrmeister2011multiwavelength, fuhrmeister2022high, howard2022mouse} placing it in the transition region between saturated and unsaturated regimes. Only at rotation periods beyond 100 days, flaring activity declines to yield coronal temperatures $<2\,$MK~\citep{wright2018stellar, foster2020corona}. For a more direct comparison, there are only few late M dwarfs beyond spectral type M6 with a spectrally resolved corona. Their intrinsic faintness and gradual decline in coronal emission make them hard to detect~\citep{berger2010, cook2014trends, stelzer2022first}. One example is NLTT 33370 AB, an M7 binary with a rotation period of about $3.8\,$h (close to TIC 277's $4.56\,$h and of the same spectral type) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}. Another is TRAPPIST-1~(M8, $P_{\rm rot} = 1.4\,$d~XXXX). Its corona appears cooler, with a 1.74 MK and a 9.6 MK component~\citep{wheatley2017strong}, and an emission measure weighted temperature of 5.36 MK~\citep{brown2023coronal}.

To summarize, considering the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for $>M7$ dwarfs to compare with, it shows a coronal temperature make-up common for mid-to-late M dwarf in the saturated activity regime.

%DS Tuc A, a young G dwarf, has been observed flaring with a >50 MK temperature in the flare peaks~\citep{pillitteri2022xraya}, and PMS stars can exceed 100 MK~\citep{getman2008xray, getman2021xray}. The Sun can heat the corona up to over 30 MK during flares~\citep{benz2016flare}. A flare on AD Leo (M3.5) had a peak temperature of about 37 MK~\citep{stelzer2022great}. CC Eri, a K7 dwarf, was described with a 3T model~\citep{crespo-chacon2007xray} with energies about $10^{33}\,$erg and peak temperatures of 30 MK.

\subsection{X-ray luminosity}
\label{sec:discussion:xraylum}
TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep{wright2011stellaractivityrotation,wright2018stellar}. In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs. This tentatively points toward a supersaturation of the corona at fast rotation speeds, as has been suggested before for M dwarfs~\citep{jeffries2011investigating}, but also earlier type stars~\citep{prosser1996rosat, argiroffi2016supersaturation} and even red giants~\citep{dixon2020rotationally}. None of the M dwarf studies probe below spectral type M6. At later spectral type, a decline in coronal activity is expected as the magnetosphere transitions from a stellar to a planetary one~\citep{pineda2017panchromatic}. \citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases from the full convection boundary (M3) towards later spectral types~\citep{magaudda2022firsta}, but begins to decrease again from M7 downward. In a compilation together with previous work from \citet{stelzer2012, cook2014trends, deluca2020extras, williams2014trends}, TIC~277 appears at the lower boundary of the $L_X/L_{\rm bol}$ distribution for M7 dwarfs. However, the rotation periods are unknown for most of these late M dwarfs. Sometime projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars: TRAPPIST-1 is similar to TIC~277 with $L_X/L_{\rm bol}\sim 10^{-3}$ at its slower 1.4 d rotation period.

At the same time, strong magnetic fields are found across the board in brown dwarfs~(XXXX).

If TIC~277 is indeed systematically underluminous, these results suggest that its dynamo produces a weaker magnetic field than that of other saturated M dwarfs, which results in a slightly lower X-ray luminosity. However, since the flaring activity of TIC~277 is in line with saturated activity, this conjecture currently lacks evidence from other activity indicators~\citep[see][, for supersaturation in flaring activity in fully convective M dwarfs]{ramsay2020tess}. As often, late M dwarfs are underrepresented in these studies.

% While it is not clear whether M7 falls in the stellar regime of~\citet{reiners2022magnetism} or into that latter brown dwarf realm, or has features of both, it appears at the lower bound regardless. The small number of supersaturated stars in \citet{nunez2022factory}  can only tentatively suggest that this is a common phenomenon among mid-to-late M dwarfs when they are quickly rotating.

%Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly underluminous. This could be an effect of its fast rotation, and consequent supersaturation as in \citet{magaudda2022first}, but we lack the rotation periods to confirm this.

\subsection{Flaring activity -- a young TRAPPIST-1}
\label{sec:discussion:flares}

$\log_{10}(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca. $-1.1$, consistent with other saturated late M dwarfs~\citep{medina2020flare, murray2022study}. It is similar to TRAPPIST-1's flaring rate~(XXXX), which has been investigated for its effects on the seven terrestrial planets in its orbit.

Recent observations of TRAPPIST-1 b~(XXXX) with the James Webb Space Telescope~(XXXX) find the planet mostly stripped of an atmosphere, which the authors attribute to the intense flare driven X-ray radiation and energetic particle exposure at the short orbital periods of the innermost planet. For the habitable zone planets TRAPPIST-1 e and f, simulations suggest that flares alter their chemical composition, in particular ozone, nitric oxide and hydroxide, but that the effects would be below the noise floor of JWST~\citep{chen2021persistence}. However, given the uncertain age of TRAPPIST-1~(XXXX), the temperate planets e and f may have been entirely stripped of their atmospheres in the past gigayear(s)~\citep{garraffo2017threatening}.

TIC~277's rapid rotation is indicative of a younger age than TRAPPIST, although it has not been attributed to any star cluster so far~(XXXX). A correct understanding of whatever the upcoming observations of TRAPPIST-1 e will yield -- detection or non-detection of an atmosphere -- will depend strongly on its activity history. Studies of its younger counterparts, late M dwarfs like TIC~277, are required to empirically constrain cumulative effects of atmospheric forcing of planets. If, as this work suggests, TIC~277 is typical of stars of its spectral type and rotation rate, it may represent the youth of TRAPPIST-1. And if its energetic flares commonly occur at high latitudes, as suggested by~\citet{ilin2021giant}, future models of energetic particle exposure of habitable zone planets may have to include age dependent latitudes of particle eruption to reproduce observations.

% \begin{figure}
%     \script{FIGURE_r315_comparison.py}
%     \begin{centering}
%         \includegraphics[width=\linewidth]{figures/r315_prot.png}
%         \caption{
%          R31.5 against rotation period for fully convective M dwarfs.
%         }
%         \label{fig:r315}
%     \end{centering}
% \end{figure}

\section{Summary and Conclusions}
\label{sec:summary}
We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at 81 deg latitude, shows exceptional coronal or flaring properties to produce this flare so close to the rotational pole. We obtained 36 ks of EPIC and OM observations with XMM-Newton, and studied the five Sectors of red-optical 27-day light curves provided by TESS. We found an X-ray luminosity of \LX, and a mean quiescent coronal temperature of \Tqmean, typical of rapidly rotating fully convective M dwarfs. The flaring activity is also typical of late M dwarfs in the saturated regime, with a flare rate $f(> log 10 Eflare = 31.5 erg)\approx$\ffdr.

XMM-Newton also captured a simultaneous X-ray and optical flare with its EPIC and OM instruments, respectively. The flare energy and rate of the OM flare are consistent with the flare frequency distribution of the 18 detected TESS flares, indicating that X-ray flares on TIC 277 can be observed with XMM-Newton about once every $10\,$h. The bolometric and soft-X-ray flare energies are within an order of magnitude within each other, which places them among other stars with simultaneous X-ray and optical flare detections, although, typically, the white light emission dominates over soft X-ray~\citep{guarcello2019simultaneous, kuznetsov2021stellar}.

While TIC 277's overall coronal and flaring properties do not appear uncommon, this does not mean that all activity saturated fully convective dwarfs produce high-latitude flares. TIC 277 is an extreme rapid rotator with a rotation period of only $4.56\,$h. The detected high-latitude flare in TESS~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}. This process may in turn produce a spatially structured corona that is not resolved in our observations, so that it appears typical of slower rotators on the saturated activity branch. Less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}. Localization of more flares on the surfaces of a broader range of stars will allow us to discriminate the role of dynamo and flux emergence theories from that of coronal properties in the production of active latitudes.

\section*{Acknowledgements}
 E.I. acknowledges funding from the Deutsche Luft- und Raumfahrtgesellschaft (FKZ 50 OR 2209, project \textit{X-ray Loops}). K.P. acknowledges funding from the German \textit{Leibniz-Gemeinschaft} under project number P67/2018. 
This work made use of the open source python software packages \texttt{lightkurve}\citep{lightkurvecollaboration2018lightkurve}, astropy~\citep{robitaille2013astropy}, numpy~(XXXX), pandas~\citep{reback2022pandasdev}, altaipony~\citep{ilin2021altaipony}, and XXXX. This work has made use of data from the European Space Agency (ESA) mission
{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement.
This work is based on observations obtained with XMM-Newton, an ESA science mission with instruments and contributions directly funded by ESA Member States and NASA.
\section*{Data Availability}
All data used in this study are publicly available in their respective archives.
This study used the reproducibility software \href{https://github.com/showyourwork/showyourwork}{showyourwork}
\citep{luger2021mappinga} to create all figures, format all tables, and compile the manuscript. Each figure links to the dataset stored on Zenodo (link to be added after acceptance), and to the script that produced the figure in the git repository~\href{https://github.com/ekaterinailin/xmm_for_tic277}{github.com/ekaterinailin/xmm\_for\_tic277}. The data analysis scripts can be found under \href{https://github.com/ekaterinailin/tic277}{github.com/ekaterinailin/tic277}.

\bibliography{references}

\end{document}